---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# timesfm

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of timesfm is to make
[TimesFM](https://github.com/google-research/timesfm), Google Research's
foundation model for time series, easy to use from R. You give it an
observed series and a horizon; it gives you point forecasts and forecast
deciles with **no training, no model selection, and no tuning** — the
pretrained model reads your series as context and forecasts in a single
forward pass.

It is a sibling of the [tabfm](https://github.com/mattyoreilly/tabfm)
package (zero-shot classification/regression on tabular data), built the
same way: a thin [reticulate](https://rstudio.github.io/reticulate/)
wrapper around the official Python package.

## Installation

You can install the development version of timesfm like so:

```{r}
# install.packages("remotes")
remotes::install_github("mattyoreilly/timesfm")
```

Then, one time only, install the underlying Python package into a dedicated
virtualenv:

```{r}
library(timesfm)
install_timesfm()
```

## Example

Pass a numeric vector or `ts`, oldest observation first, and a horizon of
up to 256 steps:

```{r}
library(timesfm)

fc <- timesfm(AirPassengers, horizon = 24)
fc

fc$mean                          # point forecasts
fc$quantiles[, c("q10", "q90")]  # an 80% prediction interval
```

The first call loads the pretrained weights (downloaded once from the
Hugging Face Hub and cached); after that, forecasts in the same session are
fast.

## Making it fast

The pretrained weights are cached after the first download, but one
optional line in your `~/.Renviron` (no Hugging Face account needed) makes
cold starts noticeably faster:

```
HF_HUB_OFFLINE=1
```

It loads weights straight from the local cache instead of revalidating them
against the Hugging Face Hub on each session's first forecast. Unset it
whenever you need to download weights you don't have yet.

## Limitations

* Forecasts require a working Python installation; `install_timesfm()`
  sets everything up via reticulate.
* Univariate series only, context capped at the most recent 1024 points,
  horizon capped at 256 steps.
* TimesFM is zero-shot: there is nothing to tune. If you need covariates,
  hierarchical structure, or bespoke accuracy on one series, reach for
  [fable](https://fable.tidyverts.org) or
  [prophet](https://facebook.github.io/prophet/).
